R Notebook

### 2021-09-21 documentation_and_analysis/draftbook.Rmd
log ev_no2 graph complete

```{r}
library(janitor)
library(readODS)
library(httr)
library(here)
library(data.table)
library(sf)
library(leaflet)
library(leaflet.extras)
library(tidyverse)
```

@@

```{r}
# Geospatial of EV Vehicles in the UK 2021 Q1
uk_ev_map %>% 
  leaflet() %>% 
  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(fillColor = ~pal(x2021_q1),
    weight = 0.1,
    opacity = 0.9, 
    color = "black",
    fillOpacity = 0.8,
    highlightOptions = highlightOptions(color = "green", weight = 2,
                                        bringToFront = TRUE),
    label = uk_ev_map_labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
  addLegend(pal = pal, values = ~x2021_q1, opacity = 0.7, title = NULL,
            position = "bottomright")
```
# How many electric vehicles are on the road across the UK by LA?

```{r}
# Reading in and skipping first 5 rows
uk_ev_postcode <- read_ods(here("raw_data/ev_by_postcode.ods"), sheet = 2, skip = 6) %>% 
  rename("postcode" = "Postcode District2")
```

```{r}
# loading in shape file 
uk_shape_file_postcode <- st_read(here("raw_data/postcode_shape/EX_Sample.shp")) %>%
  clean_names() %>% 
  st_simplify(dTolerance = 1000) %>%
  st_transform("+proj=longlat +datum=WGS84") #%>% 
#  select(lad19cd, long, lat, geometry) 
```

```{r}
# Joining uk_ev + shape file 
uk_ev_map_postcode <- uk_ev_postcode %>% 
  left_join(uk_shape_file_postcode, by = c("ons_la_code_apr_2019" = "lad19cd")) %>% 
  drop_na() %>% 
  mutate(across(c(x2021_q1:x2011_q4), as.numeric)) %>% 
  st_as_sf()
```

@@

```{r}
uk_ev_longer %>% 
  ggplot() +
  aes(x = year, y = no_of_ev) +
  geom_col(fill = "lawngreen") +
  scale_x_continuous(breaks = c(2011:2020)) +
  scale_y_continuous(breaks = seq(0, 220000, by = 20000), limits = c(0, 220000)) +
  labs(title = "\nNumber of Electric Vehicles over time in the UK\n",
       x = "\nYear\n",
       y = "\nNumber of Electric Vehicles\n") +
  theme_minimal() 
```

@@

```{r}
bins <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
pal <- colorBin("Spectral", domain = final$no2, bins = bins, na.color = "transparent", reverse = TRUE)

leaflet() %>%
  addProviderTiles("CartoDB.Positron", options = providerTileOptions(noWrap = TRUE)) %>%
  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
  addHeatmap(data = final,
             lng = ~lon,
             lat = ~lat,
             intensity = ~no2,
             minOpacity = 0.1,
             max = 45,
             radius = 1,
             blur = 1) %>% 
  addLegend(pal = pal, values = final$no2,
                title="Average NO2 Conc")
```

@@

```{r}
no2_annual_mean_all %>% 
  ggplot() +
  aes(x = year, y = annual_mean_no2concentration_mg_m3) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 70, 10), limits = c(0, 70)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  labs(x = "\nYear\n", 
       y = "\nAnnual Mean NO2 Conc mg m3\n",
       title = "\nNO2 over time in the UK\n") 
```

@@

```{r}
forecast_1 %>%
  autoplot(forecast_no2, level = NULL) +
  ggtitle("Forecasts for NO2 Conc over time") +
  xlab("Year") +
  guides(colour = guide_legend(title = "Forecast")) +
  scale_y_continuous(limit = c(0, 65))
```

@@

```{r}
ev_charging %>% 
  ggplot() +
  aes(x = year, y = number_of_chargers, fill = charger) +
  geom_col() +
  scale_y_continuous(breaks = seq(0, 30000, 2500)) +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1)) +
  scale_fill_manual(labels = c("Rapid Charger", "Standard Charger"), values = c("#e41a1c", "#4daf4a")) +
  labs(title = "\nNumber of Electric Vehicle Chargers Growth in the UK\n",
       x = "\nYear\n",
       y = "\nNumber of Chargers\n",
       fill = "Charger Type")
```

@@

```{r}
ev_pop_100k <- ev_no2_tbl %>% 
  left_join(population, by = c("ons_la_code_apr_2019" = "code")) %>% 
  mutate(ev_per_100k = x2021_q1/pop_per_100k) 
```

```{r}
ev_pop_100k %>% 
  ggplot() +
  aes(x = ev_per_100k, y = mean_no2) +
  geom_point() +
  geom_smooth(se = TRUE)
```

@@

```{r}
ev_scale <- ev_pop_100k %>% 
  select(region_local_authority_apr_2019_3, x2021_q1, mean_no2) %>% 
  mutate_if(is.numeric, scale)
```

```{r}
ev_pop_100k_scale <- ev_pop_100k %>% 
  select(region_local_authority_apr_2019_3, ev_per_100k, mean_no2) %>% 
  mutate_if(is.numeric, scale)
```

```{r}
ev_scale %>% 
  ggplot() +
  aes(x = x2021_q1, y = mean_no2) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
ev_pop_100k_scale %>% 
  ggplot() +
  aes(x = ev_per_100k, y = mean_no2) +
  geom_point() +
  geom_smooth(method = lm)
```

@@

```{r}
new_car_by_fuel_gb <- read_ods(here("raw_data/new_car_by_fuel.ods"), skip = 7) %>% 
  clean_names() %>% 
  head(21) %>% 
  mutate(across(c(petrol:zero_emission8), as.numeric)) %>% 
  filter(petrol >= 100) %>% 
  mutate(traditional_fuel = total - alternative_fuels7 - zero_emission8) %>% 
  select(c("date", "petrol", "diesel", "alternative_fuels7", "zero_emission8")) %>% 
  pivot_longer(cols = c(petrol:zero_emission8), names_to = c("fuel_type"), values_to = c("count"))
```

```{r}
new_car_by_fuel_gb %>% 
  ggplot() +
  aes(x = date, y = count, colour = fuel_type) +
  geom_line(aes(group = fuel_type)) +
  theme_minimal() +
  labs(title = "\nNew Car Registrations by Fuel Type\n",
       x = "\nYear\n",
       y = "\nNumber of Cars ('000s)\n",
       colour = "Fuel Type") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_colour_manual(labels = c("Alternative Fuels", "Diesel", "Petrol", "Zero Emission"), values = c("#abdda4", "#d7191c", "#fdae61", "#2b83ba")) 
```

>>>>

```{r}
library(janitor)
library(readODS)
library(httr)
library(here)
library(data.table)
library(sf)
library(leaflet)
library(leaflet.extras)
library(tidyverse)
library(grDevices)
```

@@

@@

```{r}
uk_ev_longer %>% 
  ggplot() +
  aes(x = year, y = no_of_ev) +
  geom_col(fill = "#08a44c") +
  scale_x_continuous(breaks = c(2011:2020)) +
  scale_y_continuous(breaks = seq(0, 220000, by = 20000), limits = c(0, 220000)) +
  labs(title = "\nNumber of Electric Vehicles over time in the UK\n",
       x = "\nYear\n",
       y = "\nNumber of Electric Vehicles\n") +
  theme_minimal() 
```

@@

```{r}
bins <- 10
pal <- colorBin("Spectral", domain = final$no2, bins = bins, na.color = "transparent", reverse = TRUE)

leaflet() %>%
  addProviderTiles("CartoDB.Positron", options = providerTileOptions(noWrap = TRUE)) %>%
  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
  addHeatmap(data = final,
             lng = ~lon,
             lat = ~lat,
             intensity = ~no2,
             minOpacity = 0.1,
             max = 45,
             radius = 1,
             blur = 1) %>% 
  addLegend(pal = pal, values = final$no2,
                title="Average NO2 Conc in 2019")
```

@@

```{r}
no2_annual_mean_all %>% 
  ggplot() +
  aes(x = year, y = annual_mean_no2concentration_mg_m3) + 
  geom_line() +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 65, 5), limits = c(22, 62)) +
  scale_x_continuous(breaks = seq(1997, 2020, 1)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "\nYear\n", 
       y = "\nAnnual Mean NO2 Conc mg m3\n",
       title = "\nNO2 over time in the UK\n") +
png("no2_time_uk.png", units="in", width=8, height=5, res=300)
# insert ggplot code
dev.off()
```

@@

```{r}
forecast_1 %>%
  autoplot(forecast_no2, level = NULL) +
  ggtitle("Forecasts for NO2 Conc over time") +
  xlab("Year") +
  guides(colour = guide_legend(title = "Forecast")) +
    theme_minimal() +
  scale_y_continuous(breaks = seq(0, 65, 5), limits = c(0, 62)) +
  scale_x_continuous(breaks = seq(1997, 2034, 3)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "\nYear\n", 
       y = "\nAnnual Mean NO2 Conc mg m3\n",
       title = "\nNO2 over time in the UK\n") +
  scale_y_continuous(limit = c(0, 65)) +
  png("no2_time_forecast.png", units="in", width=8, height=5, res=300)
# insert ggplot code
dev.off()
```

@@

```{r}
ev_charging %>% 
  ggplot() +
  aes(x = year, y = number_of_chargers, fill = charger) +
  geom_col() +
  scale_y_continuous(breaks = seq(0, 30000, 2500)) +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1)) +
  scale_fill_manual(labels = c("Rapid Charger", "Standard Charger"), values = c("#e41a1c", "#4daf4a")) +
  labs(title = "\nNumber of Electric Vehicle Chargers Growth in the UK\n",
       x = "\nYear\n",
       y = "\nNumber of Chargers\n",
       fill = "Charger Type") +
  png("no_chargers_time_uk.png", units="in", width=8, height=5, res=300)

```

@@

```{r}
ev_pop_100k <- ev_no2_tbl %>% 
  left_join(population, by = c("ons_la_code_apr_2019" = "code")) %>% 
  mutate(ev_per_100k = x2021_q1/pop_per_100k) %>% 
  drop_na()
```

```{r}
ev_pop_100k_log <- ev_pop_100k %>% 
  mutate(ev_per_100k = log(ev_per_100k))
```

```{r}
ev_pop_100k_log %>% 
  ggplot() +
  aes(x = ev_per_100k, y = mean_no2) +
  geom_point() +
  geom_smooth(se = TRUE, method = lm)
```

@@

```{r}
ev_scale <- ev_pop_100k_log %>% 
  select(region_local_authority_apr_2019_3, x2021_q1, mean_no2) %>% 
  mutate_if(is.numeric, scale)
```

```{r}
ev_pop_100k_scale_log <- ev_pop_100k_log %>% 
  select(region_local_authority_apr_2019_3, ev_per_100k, mean_no2) %>% 
  mutate_if(is.numeric, scale)
```

```{r}
ev_scale_log %>% 
  ggplot() +
  aes(x = x2021_q1, y = mean_no2) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
ev_pop_100k_scale_log %>% 
  ggplot() +
  aes(x = ev_per_100k, y = mean_no2) +
  geom_point() +
  geom_smooth(method = lm)
```

@@

```{r}
new_car_by_fuel_gb <- read_ods(here("raw_data/new_car_by_fuel.ods"), skip = 7) %>% 
  clean_names() %>% 
  head(21) %>% 
  mutate(across(c(petrol:zero_emission8), as.numeric)) %>% 
  filter(petrol >= 100) %>% 
  mutate(traditional_fuel = total - alternative_fuels7 - zero_emission8) %>% 
  select(c("date", "petrol", "diesel", "alternative_fuels7", "zero_emission8")) %>% 
  pivot_longer(cols = c(petrol:zero_emission8), names_to = c("fuel_type"), values_to = c("count")) %>% 
  mutate(date = as.numeric(date))
```

```{r}
new_car_by_fuel_gb %>% 
  ggplot() +
  aes(x = date, y = count, colour = fuel_type) +
  geom_line(aes(group = fuel_type), size = 1.25) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 2250, 250)) +
  scale_x_continuous(limits = c(2001, 2021), breaks = seq(2001, 2020, 1)) + 
  labs(title = "\nNew Car Registrations by Fuel Type\n",
       x = "\nYear\n",
       y = "\nNumber of Cars ('000s)\n",
       colour = "Fuel Type") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
  panel.grid.minor = element_blank()) +
  scale_colour_manual(labels = c("Alternative Fuels", "Diesel", "Petrol", "Zero Emission"), values = c("#abdda4", "#d7191c", "#fdae61", "#2b83ba")) +
  theme(legend.position = "none")  +
  annotate("text", x = 2006, y = 1750, label = "Petrol", colour = "#fdae61", fontface = 2) +
  annotate("text", x = 2005, y = 1100, label = "Diesel", colour = "#d7191c", fontface = 2) +
  annotate("text", x = 2016, y = 300, label = "Alternative Fuels", colour = "#abdda4", fontface = 2) +
  annotate("text", x = 2008, y = 150, label = "Zero Emissions", colour = "#2b83ba", fontface = 2) +
annotate("text", x = 2020.60, y = 950, label = "987", colour = "#fdae61", fontface = 2) +
  annotate("text", x = 2020.60, y = 250, label = "295", colour = "#d7191c", fontface = 2) +
  annotate("text", x = 2020.60, y = 450, label = "338", colour = "#abdda4", fontface = 2) +
  annotate("text", x = 2020.60, y = 50, label = "106", colour = "#2b83ba", fontface = 2) + 
png("test.png", units="in", width=8, height=5, res=300)
# insert ggplot code
dev.off()
```

```{r}
uk_ev_top <- uk_ev %>% 
  mutate(across(c(x2021_q1:x2011_q4), as.numeric)) %>% 
  filter(region_local_authority_apr_2019_3 == "Milton Keynes" |
         region_local_authority_apr_2019_3 == "Swindon" |
         region_local_authority_apr_2019_3 == "Slough" | 
         region_local_authority_apr_2019_3 == "Leeds") %>% 
  pivot_longer(cols = c(x2021_q1:x2011_q4), names_to = c("year"), values_to = "no_of_ev") %>% 
  filter(str_detect(year, "q4")) %>% 
  mutate(year = str_sub(year, 2, -4))
```

```{r}
uk_ev_top <- uk_ev_top %>% 
  left_join(uk_shape_file, by = c("ons_la_code_apr_2019" = "lad19cd")) %>% 
  drop_na() %>% 
  st_as_sf()
```

no2_all_clean <- no2_clean %>% 
  rowid_to_column()

# Source data
xy <- no2_all_clean %>% 
  dplyr::select(x, y, rowid)

# Transformed data
pj <- project(xy, proj4string, inverse=TRUE)
latlon <- data.frame(xy, lat=pj$y, lon=pj$x)
no2_all_final <-  merge(no2_all_clean, latlon, by.x = "rowid", by.y = "rowid") %>%
  dplyr::select(lat, lon, no2, year) 
```

```{r}
uk_ev_top_2021 <- uk_ev_top %>% 
  select(ons_la_code_apr_2019, region_local_authority_apr_2019_3, year, no_of_ev)

# filter for desired polygon
#highlands <- uk_ev_stockport %>% 
 # filter(region_local_authority_apr_2019_3 == "Highland")

# st_join seems less dirty
ev_no2_all_years <- no2_all_final %>% 
  st_as_sf(coords = c("lon", "lat"), crs = st_crs(uk_ev_top_2021)) %>%
  st_join(uk_ev_top_2021, join = st_intersects, left = FALSE) 
```

```{r}
ev_no2_all_year_top <- tibble(ev_no2_all_years) %>% 
  filter(year.x == year.y) %>% 
  mutate(no2 = as.numeric(no2)) %>% 
  group_by(region_local_authority_apr_2019_3, year.x) %>% 
  mutate(mean_no2 = mean(no2))
```

```{r}
ev_no2_all_year_top %>% 
  ggplot() +
  geom_line(aes(x = year.x, y = no_of_ev, colour = region_local_authority_apr_2019_3)) +
  scale_x_continuous(breaks = c(2011:2019)) +
  scale_y_continuous(breaks = seq(0, 2750, 250)) +
  labs(title = "High EV Local Authority over Years",
       x = "Year",
       y = "Number of Electric Vehicles",
       colour = "Local Authority") +
  theme_minimal()
```

```{r}
ev_no2_all_year_top %>% 
  ggplot() +
  geom_line(aes(x = year.x, y = mean_no2, colour = region_local_authority_apr_2019_3)) +
  scale_x_continuous(breaks = c(2011:2019)) +
  scale_y_continuous(breaks = c(10:28)) +
  labs(title = "Mean NO2 per Year by Local Authority",
       x = "Year",
       y = "Mean NO2",
       colour = "Local Authority") +
  theme_minimal()
```
